Supplementary Figures

### 2020-01-29 supplementary_figures/supp6_clust.Rmd
updated supplementary figures

```{r, include=FALSE}
if(!exists('FIG_NB')){
  FIG_NB <- 0
  getFigNb <- function(increment=FALSE){
    if(increment) FIG_NB <<- FIG_NB + 1
    FIG_NB
  }
}
```

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(reshape2)
  library(dplyr)
})
theme_set(theme_cowplot(font_size = 13))
suppressMessages(devtools::load_all("../../pipComp/"))
data(datasets)
source("../misc_functions.R")
```

@@

```{r, fig.width=8, fig.height=8}
res <- readRDS("../data/clust_endSummary.rds")
res <- res[which(res$dims>2 & res$resolution %in% c(0.05, 0.15, 0.3, 0.5, 0.8, 1, 1.2)),]
res$method <- paste0( gsub("clust.","",res$clustmethod, fixed=T), 
                      " k=",res$k, 
                      ifelse( res$clustmethod=="clust.seurat", paste(" res=",res$resolution), paste0(" steps=",res$steps)) )
res$diff.nbClusters <- res$nbClusters-datasets[res$dataset,"subpopulations"]
ll <- lapply( unique(res$dataset), FUN=function(x){
  as.matrix(cast2(res[which(res$dataset==x),], formula=method~dims, fun.aggregate=mean, value.var="diff.nbClusters"))
})
names(ll) <- unique(res$dataset)
mmax <- ceiling(max(abs(sapply(ll,FUN=range))))
cf <- circlize::colorRamp2(c(-mmax, 0, mmax), colors=c("blue", "white", "red"))
chm(ll, sameScale=TRUE, scaleTitle="Difference\nto real nb\nsubpopulations\n", colors=cf, cluster_columns=FALSE, cluster_rows=FALSE)
```

### Supplementary Figure `r getFigNb()`
Difference between the number of detected clusters and the number of real subpopulations according to different numbers of dimensions and clustering paramters.

>>>>

```{r, include=FALSE}
if(!exists('FIG_NB')){
  FIG_NB <- 0
  getFigNb <- function(increment=FALSE){
    if(increment) FIG_NB <<- FIG_NB + 1
    FIG_NB
  }
}

suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(reshape2)
  library(dplyr)
  library(patchwork)
})
theme_set(theme_cowplot(font_size = 12))
devtools::load_all("../../pipeComp/")
source("../misc_functions.R")
```

@@

```{r, include=FALSE}
#res <- readRDS("../data/clust_endSummary.rds")
res <- readRDS("../data/results_clust_2.rds")
res <- parsePipNames(res$clustering)
res <- res[which(res$norm=="norm.seurat" & res$resolution <5),]
x <- grep("true\\.nbClusts",colnames(res),value=TRUE)
names(x) <- x <- sapply(strsplit(x," "),FUN=function(x) x[[1]])
y <- cbind(res[,c("k","steps","resolution","clustmethod","dims")],sapply(x, FUN=function(x){
  res[[paste(x,"n_clus")]] - res[[paste(x,"true.nbClusts")]]
}))
y$param <- ifelse( res$clustmethod=="clust.seurat", paste0(" res=",res$resolution), paste0(" steps=",res$steps))
ag <- aggregate(y[,names(x)], by=y[,c("clustmethod", "k","param"),drop=FALSE], FUN=mean)
agl <- cbind(ag[rep(1:nrow(ag), 9),1:3], dataset=rep(colnames(ag)[4:ncol(ag)], each=nrow(ag)), delta=as.numeric(as.matrix(ag[,4:ncol(ag)])))
agl$k <- factor(agl$k, as.character(sort(unique(agl$k))))
x <- sapply(unique(strsplit(ag$param, "=")), FUN=function(x) as.numeric(x[[2]]) )
names(x) <- unique(ag$param)
agl$param <- factor(agl$param, names(sort(x)))

ssq <- trans_new("ssq", function(x) sign(x)*sqrt(abs(x)), function(x) sign(x)*x^2)
#slg <- trans_new("slg", function(x) sign(x)*log1p(abs(x)), function(x) sign(x)*(exp(x)-1))
pf <- function(x) ggplot(x, aes(k, param, fill=delta)) + geom_tile() + 
  facet_wrap(~dataset, nrow = 1) + labs(x="", fill="Difference to real number of clusters") + 
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank(), axis.line.x = element_blank()) +
  scale_fill_gradient2(low="darkblue", mid="white", high="red", na.value = "red", trans=ssq, limit=c(-7,20), breaks=c(-7,-2,0,2,7,20)) + 
  theme(legend.position = "none")
```

```{r, fig.width=8, fig.height=8}
(pf(agl[which(agl$clustmethod=="clust.seurat"),]) + ylab("Seurat") + theme(legend.position = "top")) / 
    (pf(agl[which(agl$clustmethod=="clust.scran"),]) + ylab("scran walktrap")) /
    (pf(agl[which(agl$clustmethod=="clust.scran.fg"),]) + ylab("scran fast_greedy")) /
    (ggplot(agl, aes(k)) + theme(axis.line.y=element_blank(), plot.margin = unit(rep(0,4),"pt"))) + 
    plot_layout(heights = c(10,8,8,1))
```

### Supplementary Figure `r getFigNb()`
Difference between the number of detected clusters and the number of real subpopulations according to different clustering paramters.

### 2020-01-29 supplementary_figures/supp_figures.tex
updated supplementary figures

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.5cm]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Supplementary Figures},
            pdfauthor={Pierre-Luc Germain},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{plainnat}
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Supplementary Figures}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Pierre-Luc Germain}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{18 Mai, 2019}

@@

\includegraphics{supp_figures_files/figure-latex/misclass-1.pdf} \vfill
\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-7-1.pdf}

@@

\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-8-1.pdf}
\vfill
\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-9-1.pdf}

@@

\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-10-1.pdf}

@@

\includegraphics{supp_figures_files/figure-latex/filt_more_ari-1.pdf}

@@

\textbf{A:} Adjusted rand index (ARI) of various clustering pipelines
using various combinations of filtering criteria. Only clustering
analyses with a number of clusters +/-1 from the real number of clusters
were considered. In general, filtering more does not result in more
accurate clustering. \textbf{B:} Percentage of cells filtered out per
subpopulation, using various combination of filters (see methods).

@@

\begin{verbatim}
## lm(ARI~nbClusters*dataset+norm+factor(filt.times)+factor(filt.MADs)+counts.lower+counts.higher+mt.higher+feat.lower+feat.higher+top50.lower+top50.higher+ratiodist.lower+ratiodist.higher)
\end{verbatim}

\begin{longtable}[]{@{}lrrrr@{}}
\toprule
& Estimate & Std. Error & t value & p value\tabularnewline
\midrule
\endhead
normnorm.seuratvst & 0.0069338 & 0.0003104 & 22.3415248 &
0.0000000\tabularnewline
factor(filt.times)2 & 0.0086024 & 0.0003617 & 23.7847962 &
0.0000000\tabularnewline
factor(filt.times)3 & 0.0168470 & 0.0003811 & 44.2112922 &
0.0000000\tabularnewline
factor(filt.MADs)2.5 & -0.0040702 & 0.0004199 & -9.6944075 &
0.0000000\tabularnewline
factor(filt.MADs)3 & -0.0054701 & 0.0004202 & -13.0177667 &
0.0000000\tabularnewline
factor(filt.MADs)5 & -0.0025099 & 0.0004205 & -5.9684908 &
0.0000000\tabularnewline
counts.lowerTRUE & -0.0001193 & 0.0003534 & -0.3375833 &
0.7356780\tabularnewline
counts.higherTRUE & -0.0073031 & 0.0003552 & -20.5607582 &
0.0000000\tabularnewline
mt.higherTRUE & 0.0006318 & 0.0007765 & 0.8136392 &
0.4158539\tabularnewline
feat.lowerTRUE & 0.0002074 & 0.0003494 & 0.5934348 &
0.5528917\tabularnewline
feat.higherTRUE & -0.0058113 & 0.0003506 & -16.5729653 &
0.0000000\tabularnewline
top50.lowerTRUE & -0.0000717 & 0.0003045 & -0.2354817 &
0.8138354\tabularnewline
top50.higherTRUE & -0.0000717 & 0.0003045 & -0.2354817 &
0.8138354\tabularnewline
ratiodist.lowerTRUE & -0.0009712 & 0.0003045 & -3.1895182 &
0.0014256\tabularnewline
ratiodist.higherTRUE & 0.0012337 & 0.0003045 & 4.0515679 &
0.0000509\tabularnewline
\bottomrule
\end{longtable}

@@

Linear regression analysis of ARI on various filtering criteria,
correcting for the effect of the number of clusters. Increasing the
number of distributions on which a cell must deviate in order to be
excluded (\texttt{filt.times}) tended to be positively associated with
ARI, while increasing the number of median absolute deviations required
to be an outlier (\texttt{filt.MADs}) was associated with a decrease in
ARI.

@@

\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-14-1.pdf}

@@

Impact of the set of filters on the average silhouette width of the
subpopulations (top) and on the proportion of variance in the first
components explained by clusters (bottom).

@@

\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-16-1.pdf}

@@

\includegraphics{supp_figures_files/figure-latex/norm_r2-1.pdf}

\textbf{A:} Correlation between library size (left) or the number of
detected features (right) and the residuals of a linear regression of
the first principal component on the subpopulation. \textbf{B:}
Proportion of the variance in the first 10 PCA components that is
explained by real subpopulations. The square-root of R-squared values is
plotted for ease of visualization.

\includegraphics{supp_figures_files/figure-latex/norm_silwidth-1.pdf}

Mean silhouette width of true subpopulations across the first 10
components with different normalization and scaling approaches.

@@

\begin{verbatim}
## Warning: Removed 328867 rows containing missing values (geom_path).

## Warning: Removed 328867 rows containing missing values (geom_path).
\end{verbatim}

@@

\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-22-1.pdf}

@@

@@

Proportion of the cumulative \emph{deviance} explained by real
subpulations that is retrieved through the selection. For each gene, we
compute the proportion of the variance explained by real subpopulations.
As for Supplementary Figure 15, except using deviance explained.

@@

@@

Proportion of the variance in the first 5 principal components that is
explained by real subpopulations according to different normalization
and selection procedures, selecting 1000 genes in each case.

>>>>

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.5cm]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Supplementary Figures},
            pdfauthor={Pierre-Luc Germain},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{plainnat}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Supplementary Figures}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Pierre-Luc Germain}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{29 Januar, 2020}

@@

\includegraphics{supp_figures_files/figure-latex/misclass-1.pdf} \vfill
\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-9-1.pdf}

@@

\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-10-1.pdf}
\vfill
\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-11-1.pdf}

@@

\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-12-1.pdf}

@@

\includegraphics{supp_figures_files/figure-latex/filt_more-1.pdf}

@@

Mean clustering F1 score per subpopulation, mean F1 at true number of
clusters, as well as maximum and median proportion of excluded cells per
subpopulation across various filtering strategies. Doublet removal
generally improves clustering accuracy with very mild exclusion rates,
even in datasets that do not have heterotypic doublets. Stringent
distribution-based filtering creates large cell type biases.

@@

@@

@@

\includegraphics{supp_figures_files/figure-latex/norm_nbClusters-1.pdf}

@@

Mean difference between the number of detected clusters and the number
of real subpopulations, depending on the normalization method, the
resolution and the number of dimensions used. The Kumar dataset is not
shown here due to a lack of variation in the number of clusters
detected. A rough ANOVA on
\texttt{nbClusters\textasciitilde{}dataset+norm+dims+resolution}
suggests that seuratvst (sctransform) is associated with a higher number
of clusters (p\textasciitilde{}0.002).

@@

\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-19-1.pdf}

@@

@@

\begin{verbatim}
## Warning: Removed 328867 rows containing missing values (geom_path).
\end{verbatim}

@@

@@

Proportion of the cumulative \emph{deviance} explained by real
subpulations that is retrieved through the selection. For each gene, we
compute the proportion of the variance explained by real subpopulations.
As for Supplementary Figure 12, except using deviance explained.

Clustering accuracy according to the number of genes selected using
various ranking/selection methods. \textbf{A:} Based on sctransform,
\textbf{B:} Based on standard Seurat normalization.

\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-27-1.pdf}

Mean difference between the number of detected clusters and the number
of real subpopulations, depending on the resolution and number of
dimensions used. Based on sctransform and seurat PCA. Increasing the
number of dimensions tends to decrease the number of identified
clusters, especially at resolutions around the default value.

\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-28-1.pdf}

@@

Adjusted Rand Index of clustering depending on the resolution and number
of dimensions used. Based on sctransform and seurat PCA.

@@

\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-31-1.pdf}

@@

Difference between the number of detected clusters and the number of
real subpopulations according to different clustering paramters.
